FL_abund_GAM

### 2020-09-11 analyses/FL_abund_GAM.Rmd
Refined and combined FL and BB GAMs

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>% # make year a factor
  dplyr::select(site, year, date, gdd.cum) %>% # drop unused columns
  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not samples
```

@@

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  semi_join(shared_transects) %>% # I only want BB data with corresponding floral data
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  semi_join(shared_transects, by = c("site", "date")) # only include site-dates that actually occurred

@@

# For bb-fl combos that never co-occurred, set preference = 1
pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp) %>%
  mutate(co = case_when(
    fl.abund == 0 | bb.abund == 0 ~ 0,
    fl.abund != 0 & bb.abund != 0 ~ 1)
  ) %>%
  group_by(plant.sp, bb.sp) %>%
  summarize(co.max = max(co)) %>%
  filter(co.max == 0) %>%
  mutate(preference.update = rep(1, n())) %>%
  dplyr::select(-co.max) %>%
  inner_join(fl_abund_sp)

@@

  # Add the forbidden links with preference set to 1 (neutral)
  full_join(pref_forbidden_link_fill) %>%
  mutate(preference = case_when(
    is.na(preference) ~ preference.update,
    !is.na(preference) ~ preference)
    ) %>%
  dplyr::select(-preference.update) %>%

  mutate(fl.abund.wt = fl.abund*preference) %>%
  mutate(yday = yday(date)) %>%
  group_by(site, date, yday, bb.sp) %>%
  summarize(fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
  ungroup() %>%
  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
         bb.sp = factor(bb.sp), # convert bb.sp to a factor
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year) %>%
  mutate(total.transects.year = n()) %>%
  ungroup()

@@

GAM modeling of preference-weighted abundance per bb species

### Would these models be better specified using gamm4? Would that let me handle the crossed and nested random effects better?

I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures. I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot. This does, of  course, create identifiability issues insofar as site and elevation are concurved. But I think I need to accept these problems. The seriousness of the identifiability problems depends in part, I think, on how mgcv handles random effect smooths. Does it first account for all the variation it can with the main effects, then model the residuals with the random effect? In this case, it's probably not a big problem, since the "partial" influence of site would orthogonal (right?) with that of te(elev.mean, yday).

```{r}
### A gamm4 version
weighted_fl_abundance.gamm4 <- gamm4(log(fl.abund.wt + 0.001) ~
                                       t2(yday, elev.mean, 
                                          by = bb.sp,
                                          #id = 0,
                                          bs= c("gp", "tp"), 
                                          k=c(8, 8), 
                                          m=1),
                                     random = ~(1 | site:year),
                                     data = fl_abund_wt,
                                     discrete = TRUE,
                                     family = "gaussian",
                                     select = TRUE,
                                     method = "fREML") %>% getViz(nsim = 500)

### Global model
#### My concern with using a global model is that, because I have so much higher sampling density in 2012, the global model will basically be just a 2012 model with 2010 and 2011 in the residuals, so to speak. I think this is at least partly mitigated by included a global by = year smooth, but I still don't quite understand how hierarchical effects relate to one another in a GAM. I'm also concerned that if I include the by = year smooth, it will be too concurve with my by = bb.sp smooths, and it will be hard to interpret the pattern of resource availability per species.
weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      s(site, bs = "re") +
                                        te(yday, elev.mean, 
                                         by = year,
                                                id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(5, 5), 
                                                m=2) +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

@@

weighted_fl_abundance.global2.ck <- check2D(weighted_fl_abundance.global2, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.global2.ck + l_gridCheck2D(gridFun = mean)

### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

weighted_fl_abundance.2010b <- bam(fl.abund.wt + 0.001 ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(15, 15), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "tw",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
check.gamViz(weighted_fl_abundance.2010b)

weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.2010b, allTerms = TRUE), pages = 4)

plot(sm(weighted_fl_abundance.2010, 2))
plot(sm(weighted_fl_abundance.2010b, 2))

weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2010b.ck <- check2D(weighted_fl_abundance.2010b, 
                                         x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)

### 2011
weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2011)
weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2011.ck + l_gridCheck2D(gridFun = mean)

### 2012
weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2012)
weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2012.ck + l_gridCheck2D(gridFun = mean)
```

### Tabulate plots
```{r}
bssFL_2010 <- plot(sm(weighted_fl_abundance.2010, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
bssFL_2011 <- plot(sm(weighted_fl_abundance.2011, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
bssFL_2012 <- plot(sm(weighted_fl_abundance.2012, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)

gersFL_2010 <- plot(sm(weighted_fl_abundance.2010, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
gersFL_2011 <- plot(sm(weighted_fl_abundance.2011, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL) 
gersFL_2012 <- plot(sm(weighted_fl_abundance.2012, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)

hortFL_2010 <- plot(sm(weighted_fl_abundance.2010, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
hortFL_2011 <- plot(sm(weighted_fl_abundance.2011, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
hortFL_2012 <- plot(sm(weighted_fl_abundance.2012, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)

humiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
humiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)
humiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. humilis")), x = NULL, y = NULL)

hypnFL_2010 <- plot(sm(weighted_fl_abundance.2010, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
hypnFL_2011 <- plot(sm(weighted_fl_abundance.2011, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
hypnFL_2012 <- plot(sm(weighted_fl_abundance.2012, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)

joneFL_2010 <- plot(sm(weighted_fl_abundance.2010, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
joneFL_2011 <- plot(sm(weighted_fl_abundance.2011, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
joneFL_2012 <- plot(sm(weighted_fl_abundance.2012, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)

lapiFL_2010 <- plot(sm(weighted_fl_abundance.2010, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
lapiFL_2011 <- plot(sm(weighted_fl_abundance.2011, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
lapiFL_2012 <- plot(sm(weighted_fl_abundance.2012, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)

mendFL_2010 <- plot(sm(weighted_fl_abundance.2010, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
mendFL_2011 <- plot(sm(weighted_fl_abundance.2011, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
mendFL_2012 <- plot(sm(weighted_fl_abundance.2012, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)

montFL_2010 <- plot(sm(weighted_fl_abundance.2010, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
montFL_2011 <- plot(sm(weighted_fl_abundance.2011, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
montFL_2012 <- plot(sm(weighted_fl_abundance.2012, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)

muciFL_2010 <- plot(sm(weighted_fl_abundance.2010, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
muciFL_2011 <- plot(sm(weighted_fl_abundance.2011, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
muciFL_2012 <- plot(sm(weighted_fl_abundance.2012, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)

pascFL_2010 <- plot(sm(weighted_fl_abundance.2010, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
pascFL_2011 <- plot(sm(weighted_fl_abundance.2011, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
pascFL_2012 <- plot(sm(weighted_fl_abundance.2012, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)

pratFL_2010 <- plot(sm(weighted_fl_abundance.2010, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
pratFL_2011 <- plot(sm(weighted_fl_abundance.2011, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
pratFL_2012 <- plot(sm(weighted_fl_abundance.2012, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)

psitFL_2010 <- plot(sm(weighted_fl_abundance.2010, 16)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
psitFL_2011 <- plot(sm(weighted_fl_abundance.2011, 16)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
psitFL_2012 <- plot(sm(weighted_fl_abundance.2012, 16)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)

pyreFL_2010 <- plot(sm(weighted_fl_abundance.2010, 17)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
pyreFL_2011 <- plot(sm(weighted_fl_abundance.2011, 17)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
pyreFL_2012 <- plot(sm(weighted_fl_abundance.2012, 17)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)

soroFL_2010 <- plot(sm(weighted_fl_abundance.2010, 18)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
soroFL_2011 <- plot(sm(weighted_fl_abundance.2011, 18)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
soroFL_2012 <- plot(sm(weighted_fl_abundance.2012, 18)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)

wurfFL_2010 <- plot(sm(weighted_fl_abundance.2010, 19)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
wurfFL_2011 <- plot(sm(weighted_fl_abundance.2011, 19)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
wurfFL_2012 <- plot(sm(weighted_fl_abundance.2012, 19)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  #l_points() +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)

major_sp <- gridPrint(bssFL_2010, bssFL_2011, bssFL_2012, 
                      pascFL_2010, pascFL_2011, pascFL_2012,
                      pratFL_2010, pratFL_2011, pratFL_2012,
                      soroFL_2010, soroFL_2011, soroFL_2012,
                      wurfFL_2010, wurfFL_2011, wurfFL_2012, ncol = 3)

high_elev_sp <- gridPrint(pyreFL_2010, pyreFL_2011, pyreFL_2012, 
                          mendFL_2010, mendFL_2011, mendFL_2012,
                          montFL_2010, montFL_2011, montFL_2012, ncol = 3)

minor_sp <- gridPrint(gersFL_2010, gersFL_2011, gersFL_2012,
                      hortFL_2010, hortFL_2011, hortFL_2012,
                      #humiFL_2010, humiFL_2011, humiFL_2012, # too rarely seen, I think only 3 times
                      hypnFL_2010, hypnFL_2011, hypnFL_2012,
                      joneFL_2010, joneFL_2011, joneFL_2012,
                      lapiFL_2010, lapiFL_2011, lapiFL_2012,
                      #psitFL_2010, psitFL_2011, psitFL_2012, # Let's ignore 
                      muciFL_2010, muciFL_2011, muciFL_2012, ncol = 3)
plot(major_sp)
plot(high_elev_sp)
plot(minor_sp)

```

>>>>

@@

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance, filtered to include only samples with corresponding floral survey data
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  semi_join(shared_transects) %>% # I only want BB data with corresponding floral data
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  semi_join(shared_transects, by = c("site", "date")) # only include site-dates that actually occurred

@@

# For bb-fl combos that never co-occurred, set preference = 1
pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp) %>%
  mutate(co = case_when(
    fl.abund == 0 | bb.abund == 0 ~ 0,
    fl.abund != 0 & bb.abund != 0 ~ 1)
  ) %>%
  group_by(plant.sp, bb.sp) %>%
  summarize(co.max = max(co)) %>%
  filter(co.max == 0) %>%
  mutate(preference.update = rep(1, n())) %>%
  dplyr::select(-co.max) %>%
  inner_join(fl_abund_sp) %>%
  select(-year)

@@

  # Add the forbidden links with preference set to 1 (neutral)
  full_join(pref_forbidden_link_fill) %>%
  mutate(preference = case_when(
    is.na(preference) ~ preference.update,
    !is.na(preference) ~ preference)
    ) %>%
  dplyr::select(-preference.update) %>%
  mutate(fl.abund.wt = fl.abund*preference) %>%
  mutate(yday = yday(date)) %>%
  group_by(site, date, yday, bb.sp) %>%
  summarize(fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
  ungroup() %>%
  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
         bb.sp = factor(bb.sp), # convert bb.sp to a factor
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year) %>%
  mutate(total.transects.year = n()) %>%
  ungroup()

@@

# GAM modeling of preference-weighted abundance per bb species
## Global model
```{r}
#### My concern with using a global model is that, because I have so much higher sampling density in 2012, the global model will basically be just a 2012 model with 2010 and 2011 in the residuals, so to speak. I think this is at least partly mitigated by included a global by = year smooth, but I still don't quite understand how hierarchical effects relate to one another in a GAM. I'm also concerned that if I include the by = year smooth, it will be too concurve with my by = bb.sp smooths, and it will be hard to interpret the pattern of resource availability per species.
weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      #s(site, bs = "re") +
                                        te(yday, elev.mean, 
                                         by = year,
                                                id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(5, 5), 
                                                m=2) +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

@@

weighted_fl_abundance.global2.ck <- check2D(weighted_fl_abundance.global2, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.global2.ck + l_gridCheck2D(gridFun = mean)
```

## Annual models
```{r}
### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")

### 2011
weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2011)
weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
                                         x1 = "yday", x2 = "elev.mean")

### 2012
weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2012)
weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
                                         x1 = "yday", x2 = "elev.mean")
```

# Extract predictions
```{r}
# Function to build data set over which to generate predictions
newdata_build <- function(dat, y) {

  ## yday min/max
  min_yday <- min(filter(dat, year == y)$yday)
  max_yday <- max(filter(dat, year == y)$yday)

  ## elev min.max
  min_elev <- min(filter(dat, year == y)$elev.mean)
  max_elev <- max(filter(dat, year == y)$elev.mean)

  ## species list
  spp <- dat %>%
    filter(year == y) %>%
    select(bb.sp) %>%
    unique() %>%
    na.omit() ##### NEED TO FIGURE OUT WHERE THE NA COMES FROM!

  # sequence of 50 points along the yday and elev ranges, respectively
  seq_yday <- seq(min_yday, max_yday, length.out = 50)
  seq_elev <- seq(min_elev, max_elev, length.out = 50)

  newdata <- tibble(seq_yday, seq_elev) %>% # combine seqs into data frame
    complete(seq_yday, seq_elev) %>% # get all possible values
    crossing(spp) %>% # get all possible combos with species
    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
}

newdata2010 <- newdata_build(fl_abund_wt, 2010)
newdata2011 <- newdata_build(fl_abund_wt, 2011)
newdata2012 <- newdata_build(fl_abund_wt, 2012)

# A little wrapper to make predictions and return a tibble with desired columns 
predict_func <- function(mod, new, y) {
  predict.bam(mod, newdata = new, type = "response") %>%
    data.frame() %>%
    tibble() %>%
    bind_cols(new) %>%
    select(yday, elev.mean, bb.sp, pred = 1) %>%
    mutate(pred.backtrans = exp(pred),
           year = rep(y, n()))
}

# Generate predictions
predict2010 <- predict_func(weighted_fl_abundance.2010, newdata2010, 2010)
predict2011 <- predict_func(weighted_fl_abundance.2011, newdata2011, 2011)
predict2012 <- predict_func(weighted_fl_abundance.2012, newdata2012, 2012)
predict_pool <- bind_rows(predict2010, predict2011, predict2012)

# Subset predictions
predict_major <- filter(predict_pool, bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf"))
predict_highelev <- filter(predict_pool, bb.sp %in% c("pyre", "mend", "mont"))
predict_minor <- filter(predict_pool, bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci"))

ggplot(predict_major, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_highelev, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)

ggplot(predict_minor, aes(yday, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25, binwidth = 1) +
  scale_fill_viridis_c() +
  facet_grid(year~bb.sp)
```
